---
title: "ABC SMC 2022"
author: "Kala Studens"
date: "`r Sys.Date()`"
output: word_document
---
```{r setup, echo = FALSE, message = FALSE, warning = FALSE}
library(sf)
library(tidyverse)
library(raster)
select <- dplyr::select

rs.all.cut2 <- read.csv('data/rs_all_example.csv') %>%
  select(-c(TrajNum, MetNum, Min, ForecastHour, 
            MIXDEPTH, TERR_MSL, ID, ID2, PRESSURE, Year))
l2 <- read_sf('data/L2.dbf') %>%
  st_transform(crs = CRS('+init=epsg:6623'))

#x.mat <- l2[,grep('X201', names(l2))]
x.mat <- l2[,grep('^201', names(l2))]
x.mat <- x.mat[,2:ncol(x.mat)] 
imm.mat <- l2[,grep('Imm_201', names(l2))]
all.mat <-  x.mat*imm.mat
rast.df2 <- all.mat %>%
  st_as_sf() %>%
  st_drop_geometry() %>%
  mutate(geometry = l2$geometry) %>%
  st_as_sf()

l2.counts <- rast.df2 %>% 
  pivot_longer(!geometry, names_to = 'Year', values_to = 'L2 Count')


ext.ind <- st_bbox(rast.df2)
ext.ind.10k <- sapply(1:length(ext.ind), function(x) {
  if (substr(names(ext.ind)[x], 2, 4) == 'min') {
    floor(ext.ind[x]/10000)*10000
  }
  else {ceiling(ext.ind[x]/10000)*10000}
})
rast <- raster(xmn = ext.ind.10k['xmin'], xmx = ext.ind.10k['xmax'],
               ymn = ext.ind.10k['ymin'], ymx = ext.ind.10k['ymax'],
               resolution = c(10000, 10000)) 

rlst <- lapply(2013:2018, function(x) {
  colnm <- paste0('X', x)
  rasterize(rast.df2, rast, field = colnm, fun = 'sum')
})

names(rlst) <- 2013:2018
rstack <- stack(rlst)

## Read in start point rasters
defoliation <- raster::stack("data/sbw_defol_stack.grd")[[7:12]]
```
**OBJECTIVE**

Use Approximate Bayesian Computation with Sequential Monte 
Carlo steps to fit SBW moth flight parameters based on presumed takeoffs from
newly defoliated areas and landings in areas with high counts of new L2s.

**OBSERVED DATA**

L2 observation sites where an increase of >= 20 larvae was 
observed from one year to the next. The L2 counts are stored in a shapefile with
counts for each year (X20[..] columns), and indicators of a likely event 
(Imm20[..] columns). Multiplying these columns gives NA for no observation, 
0 for no observed event, and a real number indicating L2 count for an 
observation of a likely immigration event.

```{r L2 Counts, warning = FALSE, message = FALSE}
ggplot(data = l2.counts) +
  geom_sf(aes(size = `L2 Count`)) +
  facet_wrap(vars(Year)) +
  theme_minimal()
```


**DATA GENERATING PROCESS**

To emulate the above data, the new defoliation maps are used in conjunction with 
NOAA's atmospheric dispersal model, HYSPLIT 
(https://www.ready.noaa.gov/HYSPLIT.php).

```{r defoliation}
plot(defoliation)
```

Due to the computational cost of 
running HYSPLIT, a large dataset of trajectories was generated prior to the 
model fitting process. Trajectories (max 9hrs in duration) were generated from 
every newly defoliated pixel for 4 hours every night from June 15th 
to August 6th of each year (9pm, 11pm, 1am, 3am). HYSPLIT requires starting
altitude as an input parameter, so the trajectories were generated at 5 discrete
altitudes: 0.4, 0.6, 0.8, 1, and 1.2x the height of the boundary layer. Each 
iteration of the ABC SMC algorithm subsets trajectories from this 13GB data 
table, "rs.all.cut2".

```{r trajectories}
head(rs.all.cut2, 20)
```

To generate a year's instance of L2 data, a date is selected at random from the
possible dates in rs.all.cut2 and a parameter vector $\theta$ is sampled from
the prior. The sampled parameters include: minimum and maximum temperatures for 
takeoff, minimum windspeed for takeoff, maximum precipitation during takeoff, 
minimum dispersal altitude and minimum temperature for dispersal.

The dataset rs.all.cut2 is then filtered based on the sampled date, 
takeoff altitude and takeoff temperature. A trajectory is only kept if: it 
originates on the sampled date, it has the correct takeoff altitude, and all 
takeoff conditions are within the boundaries specified by the parameters in 
$\theta$. The endpoints of the trajectories are filtered based on the landing 
parameters for dispesal altitude and minimum temperature. As soon as the 
dispersal altitude is reached or the minimum temperature for dispersal is 
achieved, the moths are assumed to have landed and the trajectory is cut off
at the corresponding hour; the corresponding cell in the L2 raster is marked 
1 (an observed immigration event). If a trajectory does not achieve these 
boundaries within the maximum flight time (9 hours), cells at the end of these
trajectories are marked as an observed event. The cells containing no
trajectory endpoints are marked zero.

**COMPARISON OF SIMULATED TO OBSERVED DATA**
The multi-criteria acceptance threshold uses both the number of correct cells
(true positives plus true negatives divided by # of cells), and the percentage 
of the total annual L2 count (in suspected immigration event cells) that was 
covered by the simulated data. The threshold is achieved if each value is 
greater than a specified percentile (default is 0.25 for both). 

